### 2021-03-16 paper.Rmd
cleaning up a little

The focus of this review will be on visualising high-dimensional data using linear projections, in particular, as provided by the grand tour [@Asimov1985-xr; @Buja2005-cx]. The reason being that it is not feasible to adequately the very large area of visualising high-dimensions, and there have been numerous developments in tours recently.  An overview of the technique, and new modifications is provided, along with how these techniques can be used in a variety of applications, and to help understand nonlinear dimension reduction results.

@@

<!-- Original tours (Di)
Not sure how much of this wee need to go into...
- subplots trellis literature
- data pipelines (?)
- What do we mean by multivariate data visualisation?
- multivariate categorical vis, tree maps etc.
-->

@@

> figure 2 diagram illustrating a projection: Di thinks making a Huber plot (2D data - 1D projection) here with projected data on a couple of directions is ideal

# ```{r}
# PPtreeViz::Huberplot(iris[,1:2],iris[,5],PPmethod="LDA")
# ```

@@

- Stop/start/reverse (Stuart)

@@

## Manual tour (Nick)

@@

Figure \ref{fig:manual-tour} illustrates the manual tour, by radially manipulating the contribution of variable that is crucial, to explain the separation of a particular cluster. To start we first identify a basis by orthonormalising the linear discriminant. We select a variable that has a large contribution in the basis, and watch a continuous animation as this variable is rotated fully into the projection, compeletely out of the projection, and then returned to it's original contribution.

@@

- tourr (Di) 
- spinifex (Nick)

>>>>

The focus of this review will be on visualising high-dimensional numerical data using linear projections, in particular, as provided by the grand tour [@Asimov1985-xr; @Buja2005-cx]. The reason being that it is not feasible to adequately review the very large area of visualising high-dimensions, and there have been numerous developments in tours recently. An overview of the technique, and new modifications is provided, along with how these techniques can be used in a variety of applications, and to help understand nonlinear dimension reduction results.

@@

<!-- above paragraphs could be shortened into a single paragraph? -->

<!-- 
Not sure how much of this wee need to go into...
- subplots trellis literature
- data pipelines (?)
- What do we mean by multivariate data visualisation?
- multivariate categorical vis, tree maps etc.
-->

@@

> figure 2 diagram illustrating a projection: Di thinks making a Huber plot (2D data - 1D projection) here with projected data on a couple of directions is ideal
```{r, eval = FALSE}
PPtreeViz::Huberplot(iris[,1:2],iris[,5],PPmethod="LDA")
```

@@

Due to the dynamic nature of the tour, user interfaces can enhance the interpretability of the resulting visualisation. Often one will want to pause on a particularly interesting projection and return it for use in a downstream analysis or static plot.  Basic controls can be implemented so a tour display can be paused, refreshed and replayed over and over, as shown in figure \ref{fig:tour-controls}.

@@

## Manual tour

<!-- needs to be shortend -->

@@

Figure \ref{fig:manual-tour} illustrates the manual tour, by radially manipulating the contribution of variable that is crucial, to explain the separation of a particular cluster. To start we first identify a basis by orthonormalising the linear discriminant. We select a variable that has a large contribution in the basis, and watch a continuous animation as this variable is rotated fully into the projection, completely out of the projection, and then returned to it's original contribution.

@@

- tourr (Di)

- spinifex (Nick)
